Evolution of CSMA Protocols for the IEEE 802.11 Standard

### 2013-01-23 paper.tex
I should consider the special case a=R
new csma/ca algorithm
adjusted to Boris suggestions for algorithm
new version of algorithms ready
brief description of the algorithms

@@

>>>>

%\usepackage[linesnumbered,noend]{algorithm2e}
\usepackage[linesnumbered,vlined]{algorithm2e}

@@

\begin{algorithm}
\While{True}
{
  %\tcc{Initialize retransmission attempts.}
  $r \leftarrow 0$ \;
  %\tcc{Initialize backoff stage.}
  $s \leftarrow 0$ \;
  %\tcc{Initialize backoff counter.}
  $b \leftarrow \mathcal{U}[0,2^s\rm{CW}_{min}-1]$\;
  \While{there is a packet to transmit}{
    %\tcc{Initialize $a$.}
    \Repeat{($r = R$) or (success)}{
      %\tcc{First, backoff.}
      \While{$b>0$}{
        wait 1 slot \;
        $b \leftarrow b-1$ \;
      }
      Attempt transmission \;
      \eIf{success}{
        %\tcc{Random backoff.}
        $r \leftarrow 0$ \;
        $s \leftarrow 0$ \;
        $b \leftarrow \mathcal{U}[0,2^{s}\rm{CW}_{min}-1]$\;
      }
      {
        %\tcc{Random backoff.}
        $r \leftarrow r+1$ \;
        $s \leftarrow \min r(s+1,S)$ \;
        $b \leftarrow \mathcal{U}[0, 2^s \rm{CW}_{min} -1]$\;
      }
    }
    \If{r=R}
    {
      Discard packet\;
      $r \leftarrow 0$ \;
      $s \leftarrow 0$ \;
      $b \leftarrow \mathcal{U}[0,2^s \rm{CW}_{min}-1]$\;
    }
  }
}
\caption{CSMA/CA}
\label{alg:CSMA_CA}
\end{algorithm}

\begin{algorithm}
\While{True}
{
  %\tcc{Initialize retransmission attempts.}
  $r \leftarrow 0$ \;
  %\tcc{Initialize backoff stage.}
  $s \leftarrow 0$ \;
  %\tcc{Initialize backoff counter.}
  $b \leftarrow \mathcal{U}[0,2^s\rm{CW}_{min}-1]$\;
  \While{there is a packet to transmit}{
    %\tcc{Initialize $a$.}
    \Repeat{($r = R$) or (success)}{
      %\tcc{First, backoff.}
      \While{$b>0$}{
        wait 1 slot \;
        $b \leftarrow b-1$ \;
      }
      Attempt transmission \;
      \eIf{success}{
        %\tcc{Random backoff.}
        $r \leftarrow 0$ \;
        $s \leftarrow 0$ \;
        $b \leftarrow (2^{s}\rm{CW}_{min})/2$\;
      }
      {
        %\tcc{Random backoff.}
        $r \leftarrow r+1$ \;
        $s \leftarrow \min r(s+1,S)$ \;
        $b \leftarrow \mathcal{U}[0, 2^s  \rm{CW}_{min} -1]$\;
      }
    }
    \If{r=R}
    {
      Discard packet\;
      $r \leftarrow 0$ \;
      $s \leftarrow 0$ \;
      $b \leftarrow \mathcal{U}[0, 2^s \rm{CW}_{min}-1]$\;
    }
  }
}
\caption{CSMA/ECA}
\label{alg:CSMA_ECA}
\end{algorithm}

\begin{algorithm}
\While{True}
{
  %\tcc{Initialize retransmission attempts.}
  $r \leftarrow 0$ \;
  %\tcc{Initialize backoff stage.}
  $s \leftarrow 0$ \;
  %\tcc{Initialize backoff counter.}
  $b \leftarrow \mathcal{U}[0,2^s\rm{CW}_{min}-1]$\;
  \While{there is a packet to transmit}{
    %\tcc{Initialize $a$.}
    \Repeat{($r = R$) or (success)}{
      %\tcc{First, backoff.}
      \While{$b>0$}{
        wait 1 slot \;
        $b \leftarrow b-1$ \;
      }
      Attempt transmission \;
      \eIf{success}{
        %\tcc{Random backoff.}
        $r \leftarrow 0$ \;
        %$s \leftarrow 0$ \;
        $b \leftarrow (2^{s}\rm{CW}_{min})/2$\;
      }
      {
        %\tcc{Random backoff.}
        $r \leftarrow r+1$ \;
        $s \leftarrow \min r(s+1,S)$ \;
        $b \leftarrow \mathcal{U}[0, 2^s  \rm{CW}_{min} -1]$\;
      }
    }
    \If{r=R}
    {
      Discard packet\;
      $r \leftarrow 0$ \;
      %$s \leftarrow 0$ \;
      $b \leftarrow \mathcal{U}[0, 2^s \rm{CW}_{min}-1]$\;
    }
  }
}
\caption{CSMA/ECA with hysteresis}
\label{alg:CSMA_ECA_hist}
\end{algorithm}

\begin{algorithm}
\While{True}
{
  %\tcc{Initialize retransmission attempts.}
  $r \leftarrow 0$ \;
  %\tcc{Initialize backoff stage.}
  $s \leftarrow 0$ \;
  %\tcc{Initialize backoff counter.}
  $b \leftarrow \mathcal{U}[0,2^s\rm{CW}_{min}-1]$\;
  \While{there is a packet to transmit}{
    %\tcc{Initialize $a$.}
    \Repeat{($r = R$) or (success)}{
      %\tcc{First, backoff.}
      \While{$b>0$}{
        wait 1 slot \;
        $b \leftarrow b-1$ \;
      }
      Attempt transmission of $2^s$ packets \;
      \eIf{success}{
        %\tcc{Random backoff.}
        $r \leftarrow 0$ \;
        %$s \leftarrow 0$ \;
        $b \leftarrow (2^{s}\rm{CW}_{min})/2$\;
      }
      {
        %\tcc{Random backoff.}
        $r \leftarrow r+1$ \;
        $s \leftarrow \min r(s+1,S)$ \;
        $b \leftarrow \mathcal{U}[0, 2^s  \rm{CW}_{min} -1]$\;
      }
    }
    \If{r=R}
    {
      Discard packet\;
      $r \leftarrow 0$ \;
      %$s \leftarrow 0$ \;
      $b \leftarrow \mathcal{U}[0, 2^s \rm{CW}_{min}-1]$\;
    }
  }
}
\caption{CSMA/ECA with hysteresis and fair-share}
\label{alg:CSMA_ECA_hist_fair}
\end{algorithm}

Algorithm \ref{alg:CSMA_CA} describes the CSMA/CA algorithm that is used in current networks.
When a station joins the contention it initilalizes the retry attempt counter $r$ and the backoff stage $s$ to zero.
The backoff counter $b$ is initialized using a uniform random distribution and a contention window.
After each collision, the retry attempt counter and the backoff stage counter are incremented.
As a consecuence of the incremented backoff stage, a larger contention window is used.
Note that there is maximum backoff stage $S$ and a maximum retry limit $R$.
When the maximum retry  limit is reached the packet is discarded, $r$ and $s$ are reset anb a new value for $b$ is computed.
Similarly, if the packet is successfully transmitted, the $s$ and $r$ counters are reset and a new random backoff counter is computed.

Algorithm \ref{alg:CSMA_ECA} describes CSMA/ECA in which a deterministic backoff is used after successful transmission.
The only change with respect to CSMA/CA is in line 14, where the determimistic assignment takes place.

Hysteresis is included in \ref{alg:CSMA_ECA_hist}.
Adding hysteresis is as simple as removing lines 13 and 23 from Algoritm \ref{alg:CSMA_ECA}.

Finally, fair-share is implemented in \ref{alg:CSMA_ECA_hist_fair}.
Line 10 of \ref{alg:CSMA_ECA_hist} is modified to increas the number of transmitted packets from one to $2^s$.
